---
title: HMEC 240L HGF figures
output:
  html_document:
    df_print: paged
---


```{r "setup", include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)

library(tidyverse)
library(scales)
library(RColorBrewer)
library(ComplexHeatmap)

write_pdfs <- TRUE
write_csvs <- TRUE

```


```{r loadData}

study <- "hmec240l_ss4"
fileName <- "hmec240l_ss4_rr_RUV_level3.tsv"

l3 <- read_csv(paste0("../Data/",fileName), progress = FALSE) %>%
  filter(!Ligand %in% c("KNG1|HMW", "THPO|1", "LYVE1", "JAG2|Long", "DLL1|1", "AREG"),
         !ECMp %in% c("gelatin","air", "NID1|1"))

validation_data <- read_csv("../Data/IF_validation_results_20200415.csv",
                            col_types = cols(Cells = col_double(),
                                             ECM = col_character(),
                                             Ligand = col_character(),
                                             K19_LEP = col_double(),
                                             Experiment = col_character())) %>%
  rename(ECMp = ECM) %>%
  mutate(ECMp = str_replace(ECMp, "C1", "COL1"),
         Ligand = str_replace(Ligand, "Wnt", "WNT")) %>%
  filter(ECMp == "COL1",
         Cells == "240") %>%
  mutate(Experiment = str_replace(Experiment, "20180927", "Exp_1"),
         Experiment = str_replace(Experiment, "20181108", "Exp_2"),
         Experiment = str_replace(Experiment, "20190110", "Exp_3"),
         Experiment = str_replace(Experiment, "20190125", "Exp_4"),
         Experiment = str_replace(Experiment, "20190308", "Exp_5"),
         Experiment = str_replace(Experiment, "20190408", "Exp_6"),
         Experiment = str_replace(Experiment, "20200410", "Exp_7"))

create_pdfs <- TRUE

lep_color <-"#72ac4c"
mep_color <- "#fd7f7c"
```

input data is from `r paste0("Data/",fileName)`

```{r rankedLists}
#clean up the signals, group by MEP, median summarise, rank by each signal, 
#calculate the rank products of the raw and normalized signals
sem <- function(x){
  x_sem <- sd(x)/sqrt(length(x))
  return(x_sem)
}

df_MEPs <- l3 %>%
  filter(!is.na(Cytoplasm_PA_Gated_KRT19PositiveProportion),
         !ECMp %in% c("gelatin","air", "NID1|1")) %>%
  select(Cytoplasm_PA_Gated_KRT19PositiveProportionNorm,
         ECMp, Ligand) %>%
  group_by(ECMp, Ligand)%>%
  summarise(Cytoplasm_PA_Gated_KRT19PositiveProportionNorm_sem = sem(Cytoplasm_PA_Gated_KRT19PositiveProportionNorm),
            Cytoplasm_PA_Gated_KRT19PositiveProportionNorm = median(Cytoplasm_PA_Gated_KRT19PositiveProportionNorm)) %>%
  mutate(MEP=paste(ECMp,Ligand,sep = "_"))%>%
  ungroup()

#Start with the MEP summarized values, filter to the COL1 spots
df_ligands <- df_MEPs %>%
  filter(ECMp=="COL1")

#Start with the MEP summarized values, filter to the FBS-paired MEPs
df_ECMp <- df_MEPs %>%
  filter(Ligand=="FBS")

```

<br>

#### Overview
The MEMA tech paper focuses on experiment design and data normalizations. A biologically-driven finding will help validate the methods and make for a compelling story. There is a need for luminal cells in other assays and we hypothesize that MEMAs can be used to define microenvironments that promote their growth.  

The HMEC240L SS4 assay has DNA(DAPI), basal cell type (KRT5), luminal cell type (KRT19 ) and active S phase (EdU) biomarkers.  The cell counts come from the segmenting and counting the nuclei based on the DAPI signal. The mean intensity of the KRT19 stain in the cytoplasm shows a bimodal distribution that is easily gated into KRT19 high/low populations. Since each ME spot forms its own population, we can calculate the proportion of cells that are in each gate.  



```{r, fig.width=5.5, fig.height=5}
validation_ligands_depleted <- c("BMP2", "CXCL1")
validation_ligands_middle <- c("WNT3A|1", "PDGFB|1", "FGF2|3", "TGFB1||Cterminus")
validation_ligands_enriched <- c("NRG1|1", "HGF|1", "TNFSF11|1")
df_ligands_extremes <- df_ligands %>%
  mutate(enriched_Raw = Ligand %in% validation_ligands_enriched,
         depleted_Raw =  Ligand %in% validation_ligands_depleted,
         Raw_Group = "middle",
         enriched_Norm = Ligand %in% validation_ligands_enriched,
         depleted_Norm =  Ligand %in% validation_ligands_depleted,
         Norm_Group = "middle") %>%
  filter(ECMp == "COL1",
         !Ligand %in% c("TGFB1||LAP","NRG1|10","NRG1|6", "CXCL12|Beta" )) %>%
  mutate(Ligand = str_remove(Ligand, "[|].*"),
         Ligand = str_replace(Ligand, "TGFB1", "TGFB"),
         Ligand = str_replace(Ligand, "TNFSF11", "TNFS11"),
         Ligand = str_replace(Ligand, "WNT3A", "WNT3"),
         Ligand = str_replace(Ligand, "FBS", "Control")) 

df_ligands_extremes$Raw_Group[df_ligands_extremes$enriched_Raw]<- "enriched"
df_ligands_extremes$Raw_Group[df_ligands_extremes$depleted_Raw]<- "depleted"
df_ligands_extremes$Norm_Group[df_ligands_extremes$enriched_Norm]<- "enriched"
df_ligands_extremes$Norm_Group[df_ligands_extremes$depleted_Norm]<- "depleted"

```



```{r protein_info}

prep_protein_atlas_annotations <- function(x){
  res <- x[!is.na(x)]  %>%
    str_split(pattern = ",") %>%
    unlist() %>%
    tibble(annot = .) %>%
    mutate(annot = str_remove(annot, "^[ ]")) 
}
######end of functions

proteinatlas <- read_delim("../Data//proteinatlas.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

protein_annotations <- l3 %>%
  select(Ligand, ECMp) %>%
  pivot_longer(cols = everything(), names_to = "Type", values_to = "Protein") %>%
  distinct() %>%
  mutate(Gene = str_remove_all(Protein, "[|<>].*"),
         Gene = case_when(Gene == "HA" ~"HAS1",
                          Gene == "COL1" ~"COL1A1",
                          Gene == "Wnt5a" ~"WNT5A",
                          Gene == "COL4" ~ "COL4A1",
                          Gene == "PDGFAB" ~ "PDGFA",
                          Gene == "Laminin3B32"  ~"LAMA1",
                          TRUE ~ Gene)) %>%
  left_join(proteinatlas, by = c("Gene" = "Gene")) %>%
  select(Type, Protein, Gene, 'Gene synonym', Ensembl, 'Gene description', Uniprot, 'Protein class', 'Biological process', 'Molecular function', 'Disease involvement', matches('RNA cancer'), matches('RNA tissue'))

#Get molecular function for each ME protein
molecular_functions <- protein_annotations %>%
  janitor::clean_names() %>%
  select(protein, type, molecular_function) %>%
  mutate(Cytokine = str_detect(molecular_function, "Cytokine"),
         Growth_factor = str_detect(molecular_function, "Growth factor"),
         Growth_factor = str_detect(molecular_function, "Growth factor"),
         Developmental_protein = str_detect(molecular_function, "Developmental protein"),
         Mitogen = str_detect(molecular_function, "Mitogen"))

protein_class <- prep_protein_atlas_annotations(protein_annotations$`Protein class`) %>%
  count(annot, sort = TRUE) %>%
  mutate(Protein_class = fct_inorder(annot))

p_class <- ggplot(protein_class, aes(x = Protein_class, y = n)) +
  geom_col() +
  labs(title = "MEMA ligand and ECM protein classes",
       subtitle = "source: The Human Protein Atlas",
       x = "Protein class",
       y = "Count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
p_class

biological_process <- prep_protein_atlas_annotations(protein_annotations$`Biological process`) %>%
  count(annot, sort = TRUE) %>%
  mutate(Biological_process = fct_inorder(annot))

p_bio <- ggplot(biological_process, aes(x = Biological_process, y = n)) +
  geom_col() +
  labs(title = "MEMA ligand and ECM biological process",
       subtitle = "source: The Human Protein Atlas",
       x = "Biological process",
       y = "Count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
p_bio

molecular_function <- prep_protein_atlas_annotations(protein_annotations$`Molecular function`) %>%
  count(annot, sort = TRUE) %>%
  mutate(Molecular_function = fct_inorder(annot))

p_mol <- ggplot(molecular_function, aes(x = Molecular_function, y = n)) +
  geom_col() +
  labs(title = "MEMA ligand and ECM molecular function",
       subtitle = "source: The Human Protein Atlas",
       x = "Molecular function",
       y = "Count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
p_mol

if(create_pdfs){
  pdf("../plots/HMEC_HGF_protein_annotations.pdf")
  print(p_mol)
  print(p_class)
  print(p_bio)
  res <- dev.off()
}
```

#### Method
Download annotations from The Human Protein Atlas  https://www.proteinatlas.org/about/download  
Merge annotations with ligands and ECM proteins used in the MEMA  
~6 integrin family proteins did not match annotations and need to be furthered processed  
Count annotations associated with all of the proteins and display as bar graphs  
The full set of THPA annotations are `r colnames(proteinatlas)`  



#### Luminal proportion  

We can use the gated KRT19 signal in the cytoplasm and the number of DAPI stained nuclei to calculate the proportion of luminal cells at the end of the assay. We can then compare results from the raw and normalized values.  

All ligand values are from COL1 MEPs.  

Bar charts show median values with standard error of the mean.  

```{r maximalKRT19+}
  
p <- ggplot(df_ligands_extremes, aes(x=reorder(Ligand, Cytoplasm_PA_Gated_KRT19PositiveProportionNorm, FUN=median),
                    y=Cytoplasm_PA_Gated_KRT19PositiveProportionNorm,
                    fill = Norm_Group,
                    ymin = Cytoplasm_PA_Gated_KRT19PositiveProportionNorm - Cytoplasm_PA_Gated_KRT19PositiveProportionNorm_sem,
                    ymax = Cytoplasm_PA_Gated_KRT19PositiveProportionNorm + Cytoplasm_PA_Gated_KRT19PositiveProportionNorm_sem))+
  geom_col(stat = "identity")+
    geom_errorbar(alpha = .6) +
  scale_fill_manual(values = c("enriched"= lep_color, "middle" = "gray", "depleted" = mep_color)) +
  labs(x="Ligand",
       y="Luminal proportion",
       fill = "Group",
       title="Normalized luminal proportion by ligand (COL1 spots)")  +
  theme_bw() +
  theme(axis.text.x = element_text(size = rel(.9),
                                   angle=90,
                                   color = "black"))
p

if(create_pdfs){
  pdf("../plots/MEMA_norm_ligand_bars.pdf", height = 6, useDingbats = FALSE)
  print(p)
  dev_off_res <- dev.off()
}

####
##The MEMA experiments predicted that certain factors: osteomodulin (OMD), low molecular weight hyaluronan (LMWHA), and osteopontin (SPP1), would synergize with HGF to further amplify the K19+ cells, whereas HGF combined with vitronectin (VTN) would not show increase of K19+ cells. 
###

df_HGF <- df_MEPs %>%
  filter(Ligand == "HGF|1") %>%
  mutate(ECMp = case_when(ECMp == "HA<500kDa" ~ "LMWHA",
                           ECMp == "HA>500kDa" ~ "HMWHA",
                          ECMp == "SPP1|A" ~ "SPP1",
                           TRUE ~ ECMp))

p <- ggplot(df_HGF, aes(x=reorder(ECMp, Cytoplasm_PA_Gated_KRT19PositiveProportionNorm, FUN=median),
                    y=Cytoplasm_PA_Gated_KRT19PositiveProportionNorm,
                    ymin = Cytoplasm_PA_Gated_KRT19PositiveProportionNorm - Cytoplasm_PA_Gated_KRT19PositiveProportionNorm_sem,
                    ymax = Cytoplasm_PA_Gated_KRT19PositiveProportionNorm + Cytoplasm_PA_Gated_KRT19PositiveProportionNorm_sem))+
  geom_col(stat = "identity")+
    geom_errorbar(alpha = .6) +
  labs(x="ECM protein",
       y="Luminal proportion",
       fill = "Group",
       title="Normalized luminal proportion of HGF combinations")  +
  theme_bw() +
  theme(axis.text.x = element_text(size = rel(1),
                                   angle=90,
                                   color = "black"))
p

if(create_pdfs){
  pdf("../plots/MEMA_norm_HGF_bars.pdf", height = 6, useDingbats = FALSE)
  print(p)
  dev_off_res <- dev.off()
}
```


```{r compare_screen_and_validation}

ligand_order_norm <- df_ligands_extremes %>%
  arrange(Cytoplasm_PA_Gated_KRT19PositiveProportionNorm) %>%
  pull(Ligand)

show_notch <- FALSE

#### normalize within experiments
validation_data_controls <- validation_data %>%
  filter(Ligand == "Control") %>%
  group_by(Experiment, Ligand, ECMp) %>%
    summarise(K19_LEP_median = median(K19_LEP)) %>%
  ungroup() %>%
  select(Experiment, K19_LEP_median)

val_data_norm <- validation_data %>%
  left_join(validation_data_controls, by = "Experiment") %>%
  mutate(K19_LEP_norm = K19_LEP/K19_LEP_median,
         Ligand = factor(Ligand, levels = ligand_order_norm))

control_line <- val_data_norm %>%
  filter(Ligand == "Control") %>%
  pull(K19_LEP_norm) %>%
  median()

p <- ggplot(val_data_norm, aes(x = Ligand, y = K19_LEP_norm)) +
  geom_boxplot(fill = "transparent", outlier.size = 0, notch = show_notch) +
  geom_hline(yintercept = control_line, color = "blue") 
p <- p + geom_point(aes(colour = factor(Experiment)), position = position_dodge(width = 0.90), size = .8, alpha = .9) +
  labs(title = "Luminal cell proportions in validation experiments",
       subtitle = "Collagen type 1 coated wells, HMEC240L cells, batch normalized",
       y = "Proportion of luminal cells (fold change)",
       x = "ligand treatment",
       colour = "Experiment") +
  guides(color = guide_legend(override.aes = list(size=3))) +
  theme_bw() + 
  theme(axis.text.x = element_text(color = "black"))
p

if(create_pdfs){
  pdf("../plots/HGF_well_norm_boxplots.pdf", height = 5, useDingbats = FALSE)
  print(p)
  dev_off_res <- dev.off()
}

#Summarize datasets and combine
val_data_sum <- val_data_norm %>%
  group_by(Ligand, ECMp, Cells) %>%
  summarise(K19_LEP = median(K19_LEP),
            K19_LEP_norm = median(K19_LEP_norm)) %>%
  ungroup()

df <- df_ligands_extremes %>%
  filter(ECMp == "COL1",
         !Ligand %in% c("TGFB1||LAP","NRG1|10","NRG1|6" )) %>%
  mutate(Ligand = str_remove(Ligand, "[|].*"),
         Ligand = str_replace(Ligand, "TGFB1", "TGFB"),
         Ligand = str_replace(Ligand, "TNFSF11", "TNFS11"),
         Ligand = str_replace(Ligand, "WNT3A", "WNT3"),
         Ligand = str_replace(Ligand, "FBS", "Control"),
         KRT19PositivePercentNorm = 100*Cytoplasm_PA_Gated_KRT19PositiveProportionNorm) %>%
  inner_join(val_data_sum, by = c("Ligand", "ECMp"))

p <- ggplot(df, aes(x = KRT19PositivePercentNorm,
                    y = K19_LEP_norm,
                    color = Norm_Group,
                    label = Ligand)) +
  geom_point(size = 4) +
  geom_text(nudge_y = .04, size = 2, color = "black") +
  scale_color_manual(values = c("enriched"= lep_color, "middle" = "gray", "depleted" = mep_color))  + 
  labs(title = "Coated wells vs MEMAs",
              subtitle = "normalized well data",
       x = "Percent luminal cells in MEMA (normalized)",
       y = "Luminal cells in wells (fold change)") +
  theme_bw()
p

if(create_pdfs){
  pdf("../plots/HGF_well_norm_vs_MEMA_norm.pdf",  height = 6, useDingbats = FALSE)
  print(p)
  dev_off_res <- dev.off()
}

```


```{r MEMA_figure, fig.height=7, fig.width=7}

#read in raw file so that all spots are shown
df <- read_csv(paste0("../Data/",fileName), progress = FALSE) %>%
         filter(!ECMp == "air") %>%
  group_by(ArrayRow, ArrayColumn, Well) %>%
  summarise(ECMp = unique(ECMp)) %>%
  ungroup()

df_noCOL1 <- df %>%
  filter(!ECMp == "COL1")

df_COL1 <- df %>%
  filter(ECMp == "COL1")

ECMp_cols <- c(brewer.pal(10, "Paired"), brewer.pal(8, "Dark2"))
set.seed(42)
cols <- sample(ECMp_cols, length(unique(df$ECMp)),replace = TRUE)
names(cols) <- unique(df$ECMp)

p <- ggplot(df_noCOL1, aes(x = ArrayColumn, y = ArrayRow, colour = ECMp)) +
  geom_point() +
  scale_colour_manual(values = cols) +
  scale_y_reverse(breaks = c(1, 10, 20, 30, 35)) +
  coord_cartesian(x = c(1,20),
                        y = c(35,1)) +
  scale_x_continuous(breaks = c(1,10, 20)) +
  guides(colour = FALSE) +
  labs(x = "",
       y = "")+
  facet_wrap(~Well, nrow = 2) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank())
p <- p + geom_point(data = df_COL1, shape = 17, color = "black")
p

if(create_pdfs){
  pdf("../plots/General_MEMA_cartoon.pdf",useDingbats = FALSE)
  #print(p)
  print(p + theme(strip.text = element_blank(),
                  axis.text = element_blank(),
                  axis.ticks = element_blank())
    )
  res <- dev.off()
}

#Plot replicates in the original data
replicates <- df %>%
  filter(!str_detect(ECMp, "gelatin")) %>%
  select(ECMp) %>%
  group_by(ECMp) %>%
  count(name = "count") %>%
    arrange(desc(count)) %>%
  ungroup() %>%
  mutate(ECMp = fct_inorder(ECMp),
         count_per_array = count/8)
replicate_count_median <- median(replicates$count_per_array)
p <- ggplot(replicates, aes(x = ECMp, y = count_per_array)) +
  geom_col() +
  geom_hline(yintercept = replicate_count_median, color = "blue") +
  scale_y_log10() +
  labs(title = "Number of replicates per MEMA",
       subtitle = paste("median replicate count:",replicate_count_median),
       x = "ECM protein",
       y = "Replicates per array") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
p

```

```{r l3_heatmap}



#Choose measured features and remove replicate count and standard errors
#Summarize to MEP level
deselect_features <- paste("Orientation|ReplicateCount|_SE|_RR|Cells_",
                           "Spot_PA_LoessSCC$",
                           "Cytoplasm_PA_Gated_KRT19Positive_SpotCellCount$",
                           "Spot_PA_SpotCellCount$",
                           "Cytoplasm_PA_Gated_KRT19PositiveProportion$",
                           "Cytoplasm_PA_Gated_KRTClass$", "Cytoplasm_PA_Gated_KRT19Positive$",
                           "Cytoplasm_CP_AreaShape",
                           "Cytoplasm_PA_Gated_BasalLuminalNegativeProportion$",
                           sep = "|")
meps <- l3 %>%
  group_by(MEP, Ligand, ECMp) %>%
  select(matches("_CP_|_PA_")) %>%
  select(-matches(deselect_features)) %>%
  summarise_all(.funs = median) %>%
  ungroup()

#z scale within each feature
meps_clean <- meps %>%
  select(-MEP, -Ligand, -ECMp) %>%
  as.matrix() %>%
  scale() %>%
  t() %>%
  as_tibble(rownames = "feature") %>%
  drop_na() 
colnames(meps_clean) <- c("feature", meps$MEP)

dm <- meps_clean %>%
  select(-feature) %>%
    as.matrix() %>%
  rrscale::winsor(fraction = .02)

rownames(dm) <- meps_clean$feature
colnames(dm) <- meps$MEP 

ECMp <- str_remove(colnames(dm), "_.*")
ligand <- str_remove(colnames(dm), ".*_")
HGF <- str_detect(colnames(dm), "HGF")

HGF_ECMps <- case_when(colnames(dm) %in% c("HA<500kDa_HGF|1","SPP1|A_HGF|1","OMD_HGF|1") ~ "enriched",
                 colnames(dm) %in% c("VTN_HGF|1") ~ "depleted",
                 TRUE ~ "middle"
)
#annotation structure: values of enriched, depleted and middle
#Get molecular function for each ME protein
# molecular_functions <- protein_annotations %>%
#   janitor::clean_names() %>%
#   select(protein, type, molecular_function) 

ligand_molecular_function <- meps %>%
  select(Ligand, ECMp) %>%
  left_join(molecular_functions, by = c("Ligand" = "protein"))

column_ha <- HeatmapAnnotation(ECMp = ECMp,
                               ligand = ligand,
                               HGF = HGF,
                               HGF_ECMps = HGF_ECMps,
                               Cytokine = ligand_molecular_function$Cytokine,
                               Growth_factor = ligand_molecular_function$Growth_factor,
                               Developmental_protein = ligand_molecular_function$Developmental_protein,
                               Mitogen = ligand_molecular_function$Mitogen,
                               col = list(HGF = c("FALSE" = "transparent", "TRUE" = "black"),
                                          HGF_ECMps = c("depleted" = mep_color, "middle" = "transparent", "enriched"= lep_color),
                                          Cytokine = c("FALSE" = "transparent", "TRUE" = "black"),
                                          Growth_factor = c("FALSE" = "transparent", "TRUE" = "black"),
                                          Developmental_protein = c("FALSE" = "transparent", "TRUE" = "black"),
                                          Mitogen = c("FALSE" = "transparent", "TRUE" = "black")),
                               show_legend = FALSE)
meps_hm <- Heatmap(dm,
              name = "z-score",
              column_title = "HMEC 240L",
              column_title_gp = gpar(fontsize = 12),
              cluster_rows = TRUE,
              row_km = 7,
              row_gap = unit(2, "mm"),
              row_title = " ",
              show_row_names = TRUE,
              #right_annotation = row_annots,
              cluster_columns = TRUE,
              column_km = 8,
              show_column_names = FALSE,
              top_annotation = column_ha,
              row_names_gp = gpar(fontsize = 6),
              column_names_gp = gpar(fontsize = 10),
              use_raster = FALSE)
draw(meps_hm)


col_order<- column_order(meps_hm)
#convert the column order to a tibble with cluster numbers, ligands, ECMps
# foo <- imap(col_order, function(x, y){
#   tibble(row_number =x,
#          Cluster = y)
#   }) %>%
#   bind_rows() %>%
#   mutate(mep = colnames(dm)[row_number],
#          ligand = str_remove(mep, ".*_"),
#          ECMp = str_remove(mep, "_.*"))

#foo <- bind_rows(col_order, .id = "cluster")
ECMp_order <- ECMp[unlist(col_order)]
ligand_order <- ligand[unlist(col_order)]

res <- tibble(ligand = ligand_order,
              ECMp = ECMp_order) %>%
  write_csv(path = "../tables/ME_paper_hmec240l_heatmap_columns.csv")

if(create_pdfs){
  pdf("../plots/ME_paper_hmec240l_meps_feature_heatmap.pdf", width = 8, height = 7)
  print(meps_hm)
  res <- dev.off()
}

if(create_pdfs){
  # pdf("../plots/ME_paper_hmec240l_meps_feature_heatmap_legend.pdf", width = 8, height = 4)
  # print(meps_hm_legend)
  # res <- dev.off()
}

#create a heatmap with ligands paired with COL1
ligands <- dm[,str_detect(colnames(dm), "COL1")]
colnames(ligands) <- str_remove(colnames(ligands), "COL1_")

ligands_hm <- Heatmap(ligands,
              name = "value",
              column_title = "HMEC 240L ligands, paired with COL1",
              column_title_gp = gpar(fontsize = 12),
              cluster_rows = TRUE,
              row_km = 10,
              row_gap = unit(2, "mm"),
              row_title = " ",
              show_row_names = TRUE,
              #right_annotation = row_annots,
              cluster_columns = TRUE,
              column_km = 10,
              show_column_names = TRUE,
              #top_annotation = column_annots,
              row_names_gp = gpar(fontsize = 6),
              column_names_gp = gpar(fontsize = 6),
              use_raster = FALSE)
ligands_hm

if(create_pdfs){
  pdf("../plots/ME_paper_hmec240l_ligands_feature_heatmap.pdf", width = 8, height = 6)
print(ligands_hm)
res <- dev.off()
}

```

```{r OmeroIDs}

set.seed(42)
image_nbr <- 9
l3_imageIDs <- l3 %>%
  filter(MEP %in% c("COL1_BMP2", "COL1_NRG1|1", "COL1_HGF|1")) %>%
  select(Barcode,
         ECMp,
         Ligand,
         MEP,
         ImageID,
         Well) %>%
  mutate(ImageID = as.character(ImageID)) %>%
  group_by(MEP, Barcode, Well) %>%
  slice_sample(n = image_nbr) %>%
  summarise(ImageIDs = str_c(ImageID, collapse = ",")) %>%
  ungroup()

res <- write_tsv(l3_imageIDs, "../tables/hmec240l_selected_imageIDs.tsv")

```

```{r HGF_MEMA_Validation}
##The MEMA experiments predicted that certain factors: osteomodulin (OMD), low molecular weight hyaluronan (LMWHA), and osteopontin (SPP1), would synergize with HGF to further amplify the K19+ cells, whereas HGF combined with vitronectin (VTN) would not show increase of K19+ cells. 

val_data_HGF <- read_csv("../Data/IF_ECM_K19_240_20201203.csv") %>%
  filter(Ligand == "HGF") %>%
  group_by(ECM) %>%
  summarise(LEP = median(LEP)) %>%
  ungroup() 

MEMA_val_HGF <- df_HGF %>%
  mutate(ECMp = case_when(ECMp == "HA<500kDa" ~ "LMWHA",
                           ECMp == "SPP1|A" ~ "SPP1",
                           TRUE ~ ECMp)) %>%
  right_join(val_data_HGF, by = c("ECMp" = "ECM")) %>%
  drop_na()

HGF_MEMA_val_scatterplot <- ggplot(MEMA_val_HGF, aes(x = 100*Cytoplasm_PA_Gated_KRT19PositiveProportionNorm,
                    y = LEP,
                    label = ECMp)) +
  geom_point(size = 4) +
  geom_text(nudge_y = .04, size = 2, color = "black") +
  scale_color_manual(values = c("enriched"= lep_color, "middle" = "gray", "depleted" = mep_color))  + 
  labs(title = "Concordance of MEMA and coated well HGF microenvironments",
       x = "Percent luminal cells in MEMA (normalized)",
       y = "Luminal cells in wells (fold change)") +
  theme_bw()
HGF_MEMA_val_scatterplot

if(write_pdfs){
  pdf("../plots/HMEC_HGF_3C.pdf",width = 4, height = 4,useDingbats = FALSE)
  print(HGF_MEMA_val_scatterplot)
  res <- dev.off()
}
```


